---
title: "Mall Customer Segmentation - KMeans Clustering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Customer segmentation is the practice of dividing a customer base into groups of individuals that are similar in specific ways relevant to marketing, such as age, gender, and spending habits.

The goal of segmenting customers is to decide how to relate to customers in each segment in order to maximize the value of each customer to the business.
```{r}
library(cluster)
library(factoextra)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(readr)
library(cowplot)
library(scatterplot3d)

```

```{r}
#Read Data
mall_cust <- read.csv("Mall_Customers.csv")
mall_cust
```

```{r}
#Rename column
mall_cust <- rename(mall_cust,c('SpendingScore'='Spending.Score..1.100.'))
mall_cust <- rename(mall_cust,c('AnnualIncome'='Annual.Income..k..'))
mall_cust$Gender <- as.factor(mall_cust$Gender)
str(mall_cust)
mall_cust
```

```{r}
# check for missing data
sum(is.na(mall_cust))
```

```{r}
#Clustering by Age and Spending Score
df1 <- mall_cust[,c(3,5)]
set.seed(123)

# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(df1, k, nstart = 10 )$tot.withinss}
# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15
# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)
#Elbow method
fviz_nbclust(df1, kmeans, method = "wss")
```


```{r}
#Clustering by Age and Spending Score
k2 <- kmeans(df1, centers = 4, nstart = 25)
ggplot(df1, aes(x = Age, y = SpendingScore)) + 
    geom_point(stat = "identity", aes(color = as.factor(k2$cluster))) +
    scale_color_discrete(name=" ",labels=c(paste0("Cluster",1:4))) +
    ggtitle("Mall Customer Segmens", subtitle = "K-means Clustering") + ylab("Spending Score")
```
4 clusters are observed

1. High Spending Score - customers below 40 years old
2. Medium Spending Score - customers below 40 years old
3. Medium Spending Score - customers above 40 years old
4. Low Spending Score - customers across a range of ages

```{r}
#Clustering by Annual Income and Spending Score
df2 <- mall_cust[,4:5]
set.seed(123)
fviz_nbclust(df2, kmeans, method = "wss")
```

```{r}
k2 <- kmeans(df2,5, nstart = 10)

ggplot(df2, aes(x = AnnualIncome, y = SpendingScore)) + 
    geom_point(stat = "identity", aes(color = as.factor(k2$cluster))) +
    scale_color_discrete(name=" ",labels=c(paste0("Cluster",1:5))) +
    ggtitle("Mall Customer Segments", subtitle = "K-means Clustering") + ylab("Spending Score") + xlab("Annual Income")
```
5 clusters are observed

1. High Spending Score, with lower Annual Income
2. High Spending Score, with higher Annual Income
3. Medium Spending Score with medium Annual Income
4. Low Spending Score with lower Annual Income
5. Low Spending Score with higher Annual Income
```{r}
#Clustering by Age, Annual Income and Spending Score
df3 <- mall_cust[,3:5]
set.seed(123)
fviz_nbclust(df3, kmeans, method = "wss")
```


```{r}
k3 <- kmeans(df3,5, nstart = 10)
fviz_cluster(k3, geom = "point",  data = df3) + ggtitle("k = 5")

```

```{r}
with(df3, scatterplot3d(Age, SpendingScore, AnnualIncome,pch = 16, type = "p", color = k3$cluster))
```
